---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mdufa: Structured Data on FDA's Medical Device User Fee Program Performance 

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/mdufa)](https://CRAN.R-project.org/package=mdufa)
[![R-CMD-check](https://github.com/bjoleary/mdufa/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bjoleary/mdufa/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/bjoleary/mdufa/branch/main/graph/badge.svg)](https://app.codecov.io/gh/bjoleary/mdufa?branch=main)
<!-- badges: end -->

This package includes structured data from the FDA's quarterly Medical 
Device User Fee Amendments (MDUFA) programs, which are published as PDFs 
on 
[fda.gov](https://www.fda.gov/industry/medical-device-user-fee-amendments-mdufa/mdufa-reports). 
The package includes functions to scrape those PDFs to extract the data and 
includes the most recent dataset as of publication. 

Aside from some spot-checks, the data provided by the `mdufa` R 
package has had limited verification 
and may be inaccurate. Use this information at your own risk and 
verify information using the reports provided directly by FDA. 
To facilitate this, each data point provided includes information 
about its source, including a link to the FDA report from which it 
came, the relevant page number, and more. 

"Text" metrics are typically descriptions of the MDUFA performance 
goals themselves (e.g. "90% Within 320 FDA Days"). Because of the 
way these values break across lines within table cells in the PDF 
reports, they are more difficult to retrieve and may be more likely
to be incomplete or to be inaccurate. Similarly, some tables in the
reports have footnotes, and they are not included in this dataset. 

Aside from some spot-checks, the data provided by the mdufa R 
package and included in this workbook has had limited verification 
and may be inaccurate. Use this information at your own risk and 
verify information using the reports provided directly by FDA. 
To facilitate this, each data point provided includes information 
about its source, including a link to the FDA report from which it 
came, the relevant page number, and more. 

If you find a problem in this dataset, please report it 
[here](https://github.com/bjoleary/mdufa/issues).

Please carefully read the 
[license](https://github.com/bjoleary/mdufa/blob/main/LICENSE.md) 
for additional important information.

## Installation

<!-- You can install the released version of mdufa from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("mdufa") -->
<!-- ``` -->

Install the development version from 
[GitHub](https://github.com/bjoleary/mdufa) with:

``` r
# install.packages("devtools")
devtools::install_github("bjoleary/mdufa")
```

## Examples

You can access data from MDUFA 3 using  `mdufa::mdufa3`:

```{r example_1}
library(mdufa)
library(dplyr)
set.seed(1)
mdufa::mdufa3 %>%
  select(organization, program, metric_type, performance_metric, fy, value) %>%
  sample_n(5) %>%
  print()
```
Data from MDUFA 5 is available using `mdufa::mdufa5`. 
MDUFA 4 is also available using `mdufa::mdufa4` and data from 
MDUFA 2 is available using `mdufa::mdufa2`. 

Metrics can be assessed and graphed over time:

```{r example_2}
library(ggplot2)
data <-
  mdufa::mdufa4 %>%
  mdufa::filter_metrics("double") %>%
  filter(
    report_date == max(report_date),
    organization == "CDRH",
    program == "510(k)",
    performance_metric == "Average Number of Total Days to MDUFA IV Decision"
  ) %>%
  dplyr::mutate(
    fy = as.integer(fy)
  )
graph_title <-
  paste(data$performance_metric %>% unique(), "by Fiscal Year Received")
x_label <- "Fiscal Year Received"
y_label <- data$performance_metric %>% unique()
graph_caption <-
  paste0(
    "Source: ",
    data$report_description %>% unique(),
    " available at ",
    data$report_link %>% unique()
  )
graph <-
  ggplot(
    data = data,
    mapping =
      aes(
        x = fy,
        y = value
      )
  ) +
  geom_point() +
  theme_classic() +
  scale_y_continuous(limits = c(0, NA)) +
  labs(
    title = graph_title,
    y = y_label,
    x = x_label,
    caption = graph_caption
  )
plot(graph)
```

# Exporting the data

In many cases, you may wish to locally save the data from a MDUFA report into a 
commonly used format, such as excel. This can be accomplished using the data 
already included in the package: 

```{r eval=FALSE}
mdufa::export_excel(
  data = mdufa::mdufa4,
  filepath = "mdufa4_quarterly_performance.xlsx"
)
```

# Extracting data from PDF reports

You can extract data directly from MDUFA PDF reports using `extract_report()`:

```{r eval=FALSE}
# Extract from a local PDF file
data <- mdufa::extract_report(
  pdf_path = "path/to/mdufa-4_2023-11-16.pdf",
  mdufa_period = "MDUFA IV"
)

# Or from a URL
data <- mdufa::extract_report(
  pdf_path = "https://www.fda.gov/media/...",
  mdufa_period = "MDUFA V"
)
```

This function supports MDUFA III, IV, and V report formats.

# Building the datasets for yourself

Code to build the datasets for yourself can be found in
[data-raw/](data-raw/). 
